---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# basicstats

<!-- badges: start -->
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

**basicstats** makes it easier for you to produce basic summary statistics (mean, median, range, and sample size) by levels of a grouping variable in a tidy, easy-to-read summary.

## Installation

You can install the development version of `basicstats` from [GitHub](https://github.com/andrewfullerton/basicstats) with:

``` r
# install.packages("devtools")
devtools::install_github("andrewfullerton/basicstats")
```

## Using basicstats

**basicstats** contains a single functional called `basic_stats`. Simply provide some data, a numeric variable you want basic summary statistics for, and a grouping variable by which to compute those summary statistics, and voila!

### Here's how it works:

Load the package.

```{r}
library(basicstats)
```

`basic_stats` generates a tidy, easy-to-read tibble that contains the mean, median, range, and sample size by levels of your chosen grouping variable. Here's an example using `Sepal.Width` and `Species` from the `iris` dataset.

```{r}
iris |> 
  basic_stats(Sepal.Width, Species)
```

In addition to this basic functionality, `basic_stats` also dynamically and intelligently handles character to factor and data frame to tibble conversions. Here's an example passing in a data frame, `iris`, with a character variable, `Species`, as our grouping variable. The data frame is seamlessly converted to a tibble and `Species` is converted to a factor.

```{r warning=TRUE}
iris |>
  transform(Species = as.character(Species)) |>
  basic_stats(Sepal.Width, Species)
```

To get the most out of `basic_stats`, you can further modify the calculation and presentation of your summary statistics by passing additional arguments into the function via `...`. Here's an example using `trim` to remove some extreme values from the data before calculating our summary statistics.

```{r}
iris |>
  basic_stats(Sepal.Width, Species, trim = 0.3)
```

## Thanks for checking out basicstats!
