---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# basicstats

The goal of **basicstats** is to make it easier for you to get the basic summary statistics you want (mean, median, range, and sample size, of course) based on the groupings you're most interested in.

## Installation

You can install the development version of `basicstats` from [GitHub](https://github.com/andrewfullerton/basicstats) with:

``` r
devtools::install_github("andrewfullerton/basicstats")
```

## Using the `basic_stats` function

**basicstats** contains one ground-breaking, revolutionary, first-of-its-kind function called `basic_stats`. Like magic, you simply provide some data, a numeric variable you want basic summary statistics for, a grouping variable by which you want to compute those summary statistics, and voila!

### Here's how it works:

```{r}
library(basicstats)
library(tidyverse)
```

```{r}
iris |> 
  basic_stats(Sepal.Width, Species)
```

In addition to this basic functionality, `basic_stats` dynamically and intelligently handles character to factor conversions.

```{r warning=TRUE}
iris |> 
  mutate(Species = as.character(Species)) |> # convert species to character
  basic_stats(Sepal.Width, Species)
```

It also lets to modify your summary stats to your liking by passing additional arguments into the function via the `...` argument.

```{r}
iris |>
  basic_stats(Sepal.Width, Species, trim = 0.2)
```

So much power with so little code. Welcome to **basicstats**.
